---
title: "PPA Final"
author: "Kuma Luo, Wenjun Zhu, Hao Zhu"
date: "2024-05-11"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up
```{r}
library(tidycensus)
library(dplyr)
library(sf)
library(units)

```

```{r}
census_api_key("7a719566906714f32504c8b883d2c3eca45f51fe")
```

```{r}
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")
chicagoBoundary <- 
  st_read(file.path(root.dir, "/Chapter5/chicagoBoundary.geojson")) %>%  
  st_transform('ESRI:102271')
```



# Measuring Gentrification
For measuring whether a census tract is gentrified or not, our group decided to use percentage changes of median household income, property value, and population who have colledge degree or above.

## Import Data
```{r}
chicago_tracts <- st_read("data/Boundaries - Census Tracts - 2010/geo_export_a2829751-660b-4864-9839-eb6be574475f.shp")

geoids <- chicago_tracts$geoid10
```


```{r, eval=FALSE}

# Variables for educational attainment (bachelor's degree or higher)
education2009 <- c("B15002_014E", "B15002_015E", "B15002_016E", "B15002_017E", "B15002_018E",
                   "B15002_031E", "B15002_032E", "B15002_033E", "B15002_034E", "B15002_035E")

education_vars <- c("B15003_021E", "B15003_022E", "B15003_023E", "B15003_024E", "B15003_025E")

# Variable for median household income
income_var <- "B19013_001E"

# Variable for median property value (housing values)
property_var <- "B25077_001E"


# Download data for 2009 and 2019
data_2009 <- get_acs(geography = "tract",
                     variables = c(education2009, income_var, property_var),
                     year = 2009,
                     survey = "acs5",
                     state = "IL",
                     county = "Cook",
                     output = "wide") %>%
  filter(GEOID %in% geoids)


data_2019 <- get_acs(geography = "tract",
                     variables = c(education_vars, income_var, property_var),
                     year = 2019,
                     survey = "acs5",
                     state = "IL",
                     county = "Cook",
                     output = "wide") %>%
  filter(GEOID %in% geoids)

```

```{r, include=FALSE}

# Variables for educational attainment (bachelor's degree or higher)
education2009 <- c("B15002_014E", "B15002_015E", "B15002_016E", "B15002_017E", "B15002_018E",
                   "B15002_031E", "B15002_032E", "B15002_033E", "B15002_034E", "B15002_035E")

education_vars <- c("B15003_021E", "B15003_022E", "B15003_023E", "B15003_024E", "B15003_025E")

# Variable for median household income
income_var <- "B19013_001E"

# Variable for median property value (housing values)
property_var <- "B25077_001E"


# Download data for 2009 and 2019
data_2009 <- get_acs(geography = "tract",
                     variables = c(education2009, income_var, property_var),
                     year = 2009,
                     survey = "acs5",
                     state = "IL",
                     county = "Cook",
                     output = "wide") %>%
  filter(GEOID %in% geoids)


data_2019 <- get_acs(geography = "tract",
                     variables = c(education_vars, income_var, property_var),
                     year = 2019,
                     survey = "acs5",
                     state = "IL",
                     county = "Cook",
                     output = "wide") %>%
  filter(GEOID %in% geoids)

```

```{r}
process_education_data2009 <- function(data) {
  data %>%
    mutate(college_degree_or_higher = rowSums(select(., B15002_014E, B15002_015E,
                                                      B15002_016E, B15002_017E, 
                                                      B15002_018E, B15002_031E,
                                                      B15002_032E, B15002_033E, 
                                                      B15002_034E, B15002_035E), na.rm = TRUE)) %>%
    select(GEOID, college_degree_or_higher, ends_with("B19013_001E"), ends_with("B25077_001E"))
}

# Processing function for 2019 data
process_education_data2019 <- function(data) {
  data %>%
    mutate(college_degree_or_higher = rowSums(select(., B15003_021E, B15003_022E, 
                                                     B15003_023E, 
                                                     B15003_024E, B15003_025E), 
                                              na.rm = TRUE)) %>%
    select(GEOID, college_degree_or_higher, ends_with("B19013_001E"), ends_with("B25077_001E"))
}

data_2009_processed <- process_education_data2009(data_2009) %>%
    mutate(income_group = cut(B19013_001E,
                            breaks = quantile(B19013_001E, probs = c(0, 1/3, 2/3, 1), 
                                              na.rm = TRUE),
                            labels = c("Low", "Medium", "High"),
                            include.lowest = TRUE)) %>%
  filter(income_group %in% c("Low", "Medium")) %>%
  na.omit()

data_2019_processed <- process_education_data2019(data_2019)


combined_data <- left_join(data_2009_processed, data_2019_processed, by = "GEOID")


combined_data <- combined_data %>%
  mutate(
    percent_change_education = (college_degree_or_higher.y - college_degree_or_higher.x) / college_degree_or_higher.x * 100,
    percent_change_income = (B19013_001E.y - B19013_001E.x) / B19013_001E.x * 100,
    percent_change_property = (B25077_001E.y - B25077_001E.x) / B25077_001E.x * 100
  ) %>%
  na.omit()

```

## Identify Gentrified Areas

```{r}
mean(combined_data$percent_change_income) 
mean(combined_data$percent_change_property) 
mean(combined_data$percent_change_education) 

combined_data <- combined_data %>%
  mutate(gentrified = ifelse(percent_change_income > 25.45924 & 
                               percent_change_property > -11.16558 & 
                               percent_change_education > 47.80252, 1, 0)) %>%
  select(percent_change_income, percent_change_property, percent_change_education, 
         gentrified, GEOID)
```

# Create Predictive Model

## Import Data
```{r}
chicago_tracts$geoid10 <- as.character(chicago_tracts$geoid10)
combined_data$GEOID <- as.character(combined_data$GEOID)

final_data <- left_join(chicago_tracts, combined_data, by = c("geoid10" = "GEOID")) 

final_data <- final_data %>% 
  filter(!is.na(gentrified)) %>%
  select(-commarea, -countyfp10, -name10, -namelsad10, -notes, -statefp10, -tractce10)

```


Park share
```{r}
park <- st_read("data/Parks - Chicago Park District Park Boundaries (current)/geo_export_d0177a39-9cca-4395-bf70-2c21eb83726a.shp") 

park <- park %>%
  dplyr::select(geometry)


final_data <- st_make_valid(final_data)
park <- st_make_valid(park)

park_intersected_areas <- st_intersection(final_data, park)

park_intersected_areas$park_area <- st_area(park_intersected_areas) %>%
  drop_units() 

park_intersected_areas <- park_intersected_areas%>%
  select(park_area)

```

```{r}

final_data$tract_area <- st_area(final_data) %>%
  drop_units()

final_data <- st_join(final_data, park_intersected_areas, left = TRUE) 
  
final_data$Park_Share <- (final_data$park_area /final_data$tract_area) * 100

final_data$Park_Share =replace(final_data$Park_Share, is.na(final_data$Park_Share), 0)

```

Demographic Variables
```{r pressure, echo=FALSE, eval=FALSE}
library(tidycensus)
library(tidyverse)

Geoid <- final_data$geoid10

variables <- c(
 "B11001_003E", "B11001_007E", "B25024_006E", "B25024_010E", "B25044_010E", "B01001_009E", "B01001_033E","B01001_010E","B01001_034E","B01001_011E",
  "B01001_035E")

ChicagoData <- get_acs(
  geography = "tract",
  variables = variables,
  year = 2019,
  state = "IL",
  county = "Cook",
  geometry = TRUE,
  output = "wide")%>%
  filter(GEOID %in% Geoid)

ChicagoData <- ChicagoData %>%
  dplyr::select(-dplyr::contains("M"))  

ChicagoData <- ChicagoData %>% rename(
  age_30_34_m = "B01001_009E",
  age_30_34_f = "B01001_033E",
  age_35_39_m = "B01001_010E",
  age_35_39_f = "B01001_034E",
  age_40_44_m = "B01001_011E",
  age_40_44_f = "B01001_035E",
    married_with_children = "B11001_003E",  
    total_nonfamily_households = "B11001_007E",
    units_in_3_to_4_building = "B25024_006E",  
    units_in_5plus_building = "B25024_010E",
    units_with_three_cars = "B25044_010E" ) %>%
mutate(age_30_45 = age_30_34_m + age_30_34_f +
                age_35_39_m + age_35_39_f +
                age_40_44_m + age_40_44_f, units_in_3plus_building= units_in_3_to_4_building+ units_in_5plus_building)


```

```{r pressure1, echo=FALSE, include=FALSE}
library(tidycensus)
library(tidyverse)

Geoid <- final_data$geoid10

variables <- c(
 "B11001_003E", "B11001_007E", "B25024_006E", "B25024_010E", "B25044_010E", "B01001_009E", "B01001_033E","B01001_010E","B01001_034E","B01001_011E",
  "B01001_035E")

ChicagoData <- get_acs(
  geography = "tract",
  variables = variables,
  year = 2019,
  state = "IL",
  county = "Cook",
  geometry = TRUE,
  output = "wide")%>%
  filter(GEOID %in% Geoid)

ChicagoData <- ChicagoData %>%
  dplyr::select(-dplyr::contains("M"))  

ChicagoData <- ChicagoData %>% rename(
  age_30_34_m = "B01001_009E",
  age_30_34_f = "B01001_033E",
  age_35_39_m = "B01001_010E",
  age_35_39_f = "B01001_034E",
  age_40_44_m = "B01001_011E",
  age_40_44_f = "B01001_035E",
    married_with_children = "B11001_003E",  
    total_nonfamily_households = "B11001_007E",
    units_in_3_to_4_building = "B25024_006E",  
    units_in_5plus_building = "B25024_010E",
    units_with_three_cars = "B25044_010E" ) %>%
mutate(age_30_45 = age_30_34_m + age_30_34_f +
                age_35_39_m + age_35_39_f +
                age_40_44_m + age_40_44_f, units_in_3plus_building= units_in_3_to_4_building+ units_in_5plus_building)


```


```{r}
ChicagoData <- as.data.frame(ChicagoData)
final_data$geoid10 <- as.character(final_data$geoid10)
ChicagoData$GEOID <- as.character(ChicagoData$GEOID)
final_data <- final_data %>%
  left_join(ChicagoData, by = c("geoid10" = "GEOID"))%>%
  select(-geometry.y)

```


Rail Station
```{r}
stations <- st_read("data/CTA_RailStations/CTA_RailStations.shp") 
stations <- st_transform(stations, st_crs(final_data))

stopBuffer <- st_buffer(stations, 2640)

stopUnion <- st_union(st_buffer(stations, 2640))

StationBuffers <- 
  rbind(
     stopBuffer %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
     stopUnion %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))


```


```{r}
buffer <- filter(StationBuffers, Legend=="Unioned Buffer") 
  
buffer <- st_transform(buffer, st_crs(final_data))

final_data <- st_make_valid(final_data)
buffer <- st_make_valid(buffer)

buffer_intersect <- st_intersection(final_data, buffer)


```

```{r}
buffer_intersect <- buffer_intersect%>%
  select(geoid10, Legend)

buffer_intersect <- buffer_intersect %>%
  st_set_geometry(NULL)

buffer_intersect <- buffer_intersect %>%
  group_by(geoid10) %>%
  summarize(
    Count = n(),  # as before
    .groups = 'drop'  # drop groups after summarizing to avoid issues
  )


final_data <- final_data %>%
  left_join(buffer_intersect, by = ("geoid10"))
  
final_data$Count = replace(final_data$Count, is.na(final_data$Count), 0)

final_data <- final_data %>%
  mutate(Have_RailStation = ifelse(Count > 0, 1, 0)) %>%
  select(-Count)
  

```

## Visualize the Gentrified Area and Independent Variables

```{r}
ggplot() + 
  geom_sf(data=final_data, aes(fill = as.factor(gentrified))) + 
  geom_sf(data = chicagoBoundary, fill = NA, color = "black") +
  theme_minimal() +
  labs(title="Gentrification Area in Chicago") +
  theme(plot.title = element_text(hjust = 0.5),  
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
    legend.position = "right")
```




```{r}
library(tidyr)
library(ggplot2)
library(dplyr)

dat0 <- final_data %>%
  st_set_geometry(NULL)

Variables <- dat0 %>%
  pivot_longer(
    cols = c(commarea_n, Park_Share, married_with_children, total_nonfamily_households,
             units_in_3plus_building, units_with_three_cars, age_30_45, Have_RailStation),
    names_to = "variable",
    values_to = "value"
  ) %>% 
  group_by(gentrified, variable) %>%
  summarize(mean_value = mean(value, na.rm = TRUE)) %>%
  ungroup()  


ggplot(Variables, aes(x = as.factor(gentrified), y = mean_value, fill = as.factor(gentrified))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~variable, scales = "free_y") +
  scale_fill_manual(values = c("0" = "#00BFFF", "1" = "#fc2f00"),
                    labels = c("0" = "Non_Gentrified", "1" = "Gentrified"),
                    name = "Gentrification") +
  labs(
    title = "Average Values by Gentrified Status for Different Variables",
    x = "Gentrification",
    y = ""
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text.x = element_text(size = 10), # Adjust facet label size if needed
    plot.title = element_text(hjust = 0.5), # Center the main title
    axis.text.x = element_text(angle = 45, hjust = 1) # Tilt x axis labels if needed
  )


```

## Binomial Logistic Regression
```{r}
dat <- final_data %>%
  st_set_geometry(NULL) %>%  # This removes the geometry column effectively
  as.data.frame() %>%
  select(commarea_n, gentrified, Park_Share, married_with_children, total_nonfamily_households, 
         units_in_3plus_building, units_with_three_cars, age_30_45, Have_RailStation)

library(caret)

set.seed(3456)

trainIndex <- createDataPartition(dat$gentrified, p = .60,
                                  list = FALSE,
                                  times = 1)
datTrain <- dat[ trainIndex,]
datTest  <- dat[-trainIndex,]
```

```{r}
GLMModel <- glm(gentrified ~ ., family = binomial(link = "logit"), data = datTrain)

summary(GLMModel)

```

```{r}
model1 <- glm(gentrified ~ . -commarea_n, 
              family="binomial"(link="logit"), data = datTrain)

summary(model1)
```

```{r}
model2 <- glm(gentrified ~ Park_Share + married_with_children + 
         units_in_3plus_building + units_with_three_cars + age_30_45 + Have_RailStation, 
              family="binomial"(link="logit"), data = datTrain) 
summary(model2)
#final model 
```


```{r}
model3 <- glm(gentrified ~ . -commarea_n -married_with_children, 
              family="binomial"(link="logit"), data = datTrain)

summary(model3)


```
```{r}
model4 <- glm(gentrified ~ . -commarea_n -married_with_children -total_nonfamily_households, 
              family="binomial"(link="logit"), data = datTrain)

summary(model4)
```
```{r}

if (!require(lmtest)) install.packages("lmtest")


library(lmtest)

lrtest(model3, model4)
lrtest(GLMModel, model3)
lrtest(GLMModel, model2)

```
## Model Performance

```{r predict_first}
classProbs <- predict(model2, datTest, type = "response")

hist(classProbs, main = "Distribution of Predicted Gentrified Areas", xlab = "Predicted Probability", ylab = "Frequency", col = "#fc2f00")

```

##Predicted Probability Density of Gentrification in Chicago


```{r plot_preds}
plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  #panel.border=element_rect(colour="#F0F0F0"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.75),
  axis.ticks=element_blank())

testProbs <- data.frame(obs = as.numeric(datTest$gentrified),
                        pred = classProbs)


```
Confusion Metrics

```{r philly model confusion matrix}
P.predprob <- data.frame(Probs = predict(model2, type = "response", newdata = dat))

pred <- cbind.data.frame(dat, P.predprob) %>% 
  mutate(predOutcome  = as.factor(ifelse(Probs > 0.171 , "Y", "N")))

P.conf_matrix <- table(observed = pred$gentrified, predicted = pred$predOutcome)

P.conf_matrix

100 * prop.table(table(Observed = pred$gentrified, predicted = pred$predOutcome), margin = 1)
```

Cross Validation

```{r k_fold, warning = FALSE, message = FALSE}
ctrl <- trainControl(method = "cv", 
                     number = 100, 
                     p = 0.7, 
                     savePredictions = TRUE)

cvFit <- train(as.factor(gentrified) ~ .,  data = dat %>% 
                                                as.data.frame() , 
               method="glm", family="binomial",
               trControl = ctrl)

cvFit
```
Predictive Map in Chicago

```{r}
allPredictions.c <- 
  predict(model2, newdata = dat, type="response")
  
dat.c <- 
  cbind(dat,allPredictions.c) %>%
  mutate(allPredictions = round(allPredictions.c * 100), 
             predicted.gentrified = ifelse(allPredictions > 12.8, 1, 0)) 
```

```{r}
dat.c <- dat.c %>%
  mutate(geometry = st_geometry(final_data))
dat.c <- st_as_sf(dat.c, sf_column_name = "geometry")
```
Notably, the green areas, which represent gentrified tracts, show significant overlap between the two maps, demonstrating the high accuracy of our model in making predictions. Additionally, our predictions include some areas that have not yet undergone gentrification but show potential. Therefore, our model is particularly useful for policymakers to identify and preemptively address areas likely to experience gentrification trends. The maps also suggest that many tracts in the northern part of Chicago are more likely to gentrify, which we suspect is related to the influence of the city's northeast central area.

```{r}
ggplot() + 
  geom_sf(data=dat.c, aes(fill = as.factor(predicted.gentrified))) + 
  geom_sf(data = chicagoBoundary, fill = NA, color = "black") +
  theme_minimal() +
  labs(title="Predicted Gentrification Area in Chicago") +
  theme(plot.title = element_text(hjust = 0.05),  
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 8),  # Smaller x-axis text
        axis.text.y = element_text(size = 8),  # Smaller y-axis text
        legend.position = "right")

```
```{r}
 ggplot() + 
  geom_sf(data=dat.c, aes(fill = as.factor(gentrified))) + 
  geom_sf(data = chicagoBoundary, fill = NA, color = "black") +
  theme_minimal() +
  labs(title="Observed Gentrification Area in Chicago") +
  theme(plot.title = element_text(hjust = 0.05),  
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 8),  # Smaller x-axis text
        axis.text.y = element_text(size = 8),  # Smaller y-axis text
        legend.position = "right")


```




# Predict Gentrification in Philadelphia
Then we use our model to predict how gentrification might happen in Philadelphia for validation.

## Import Variables
```{r}
PhillyTracts <- st_read("data/PhillyBound/c16590ca-5adf-4332-aaec-9323b2fa7e7d2020328-1-1jurugw.pr6w.shp")

PhillyTracts <- PhillyTracts %>%
  select(GEOID10)
```


Demographic Varibales
```{r, eval=FALSE}
variablesPhilly <- c(
  "B11001_003E", "B25024_006E", "B25024_010E",
  "B25044_010E", "B01001_009E", "B01001_033E","B01001_010E","B01001_034E","B01001_011E",
  "B01001_035E"
)

PhillyData <- get_acs(
  geography = "tract",
  variables = variablesPhilly,
  year = 2019,
  state = "PA",
  county = "Philadelphia",
  geometry = TRUE,
  output = "wide")

PhillyData <- PhillyData %>%
  dplyr::select(-dplyr::contains("M"))  

PhillyData <- PhillyData %>% 
  rename(
  age_30_34_m = "B01001_009E",
  age_30_34_f = "B01001_033E",
  age_35_39_m = "B01001_010E",
  age_35_39_f = "B01001_034E",
  age_40_44_m = "B01001_011E",
  age_40_44_f = "B01001_035E", 
    married_with_children = "B11001_003E",
    units_in_3_to_4_building = "B25024_006E",  
    units_in_5plus_building = "B25024_010E",
    units_with_three_cars = "B25044_010E" ) %>%
mutate(age_30_45 = age_30_34_m + age_30_34_f +
                age_35_39_m + age_35_39_f +
                age_40_44_m + age_40_44_f, units_in_3plus_building= units_in_3_to_4_building+ units_in_5plus_building)

```

```{r, include=FALSE}
variablesPhilly <- c(
  "B11001_003E", "B25024_006E", "B25024_010E",
  "B25044_010E", "B01001_009E", "B01001_033E","B01001_010E","B01001_034E","B01001_011E",
  "B01001_035E"
)

PhillyData <- get_acs(
  geography = "tract",
  variables = variablesPhilly,
  year = 2019,
  state = "PA",
  county = "Philadelphia",
  geometry = TRUE,
  output = "wide")

PhillyData <- PhillyData %>%
  dplyr::select(-dplyr::contains("M"))  

PhillyData <- PhillyData %>% 
  rename(
  age_30_34_m = "B01001_009E",
  age_30_34_f = "B01001_033E",
  age_35_39_m = "B01001_010E",
  age_35_39_f = "B01001_034E",
  age_40_44_m = "B01001_011E",
  age_40_44_f = "B01001_035E", 
    married_with_children = "B11001_003E",
    units_in_3_to_4_building = "B25024_006E",  
    units_in_5plus_building = "B25024_010E",
    units_with_three_cars = "B25044_010E" ) %>%
mutate(age_30_45 = age_30_34_m + age_30_34_f +
                age_35_39_m + age_35_39_f +
                age_40_44_m + age_40_44_f, units_in_3plus_building= units_in_3_to_4_building+ units_in_5plus_building)

```

Spatial Data
```{r}
Phillystations <- st_read("data/Regional_Rail_Stations/Regional_Rail_Stations.shp")
Phillystations <- st_transform(Phillystations, st_crs(PhillyData))

PhillystopBuffer <- st_buffer(Phillystations, 2640)

PhillystopUnion <- st_union(st_buffer(Phillystations, 2640))

PhillyStationBuffers <- 
  rbind(
     PhillystopBuffer %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
     PhillystopUnion %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))

```
```{r}
Phillybuffer <- filter(PhillyStationBuffers, Legend=="Unioned Buffer") 
  
Phillybuffer <- st_transform(Phillybuffer, st_crs(PhillyData))

PhillyData <- st_make_valid(PhillyData)
Phillybuffer <- st_make_valid(Phillybuffer)

Phillybuffer_intersect <- st_intersection(PhillyData, Phillybuffer)


```

```{r}
Phillybuffer_intersect <- Phillybuffer_intersect%>%
  select(GEOID, Legend)

Phillybuffer_intersect <- Phillybuffer_intersect %>%
  st_set_geometry(NULL)


PhillyFinal <- PhillyData %>%
  left_join(Phillybuffer_intersect, by = ("GEOID"))
  
PhillyFinal$Legend = replace(PhillyFinal$Legend, is.na(PhillyFinal$Legend), 0)

PhillyFinal <- PhillyFinal %>%
  mutate(Have_RailStation = ifelse(Legend == "Unioned Buffer", 1, Legend)) %>%
  select(-Legend)

PhillyFinal$Have_RailStation <- as.numeric(PhillyFinal$Have_RailStation)

```

```{r}
PhillyPark <- st_read("data/PhillyFPC_PPR_Assets (1)/PhillyFPC_PPR_Assets.shp")

PhillyPark <- PhillyPark %>%
  dplyr::select(geometry)


PhillyFinal <- st_make_valid(PhillyFinal)

PhillyPark <- st_make_valid(PhillyPark)


PhillyPark <- PhillyPark %>%
  st_transform(crs = st_crs(PhillyFinal))

PhillyPark_intersected_areas <- st_intersection(PhillyFinal, PhillyPark)

PhillyPark_intersected_areas$park_area <- st_area(PhillyPark_intersected_areas) %>%
  drop_units() 

PhillyPark_intersected_areas <- PhillyPark_intersected_areas%>%
  select(park_area)

```

```{r}

PhillyFinal$tract_area <- st_area(PhillyFinal) %>%
  drop_units()

PhillyFinal1 <- st_join(PhillyFinal, PhillyPark_intersected_areas, left = TRUE)

PhillyFinal1 <- PhillyFinal1 %>%
  group_by(GEOID) %>%
  summarize(total_park_area = sum(park_area, na.rm = TRUE)) %>%
  ungroup()
  
PhillyFinal1 <- PhillyFinal1 %>%
  st_set_geometry(NULL)

PhillyFinal <- PhillyFinal %>%
  left_join(PhillyFinal1, by = ("GEOID"))
  
PhillyFinal$Park_Share <- (PhillyFinal$total_park_area /PhillyFinal$tract_area) * 100

PhillyFinal$Park_Share =replace(PhillyFinal$Park_Share, is.na(PhillyFinal$Park_Share), 0)

```

## Predicted Map

Implement Model
```{r}
dat.philly <- PhillyFinal %>%
  st_set_geometry(NULL) %>%
  as.data.frame() %>%
  select(Park_Share, married_with_children, 
         units_in_3plus_building, units_with_three_cars, age_30_45, Have_RailStation)

```


```{r}
allPredictions <- 
  predict(model2, newdata = dat.philly, type="response")
  
dat.philly <- 
  cbind(dat.philly,allPredictions) %>%
  mutate(allPredictions = round(allPredictions * 100), 
             gentrified = ifelse(allPredictions > 12.8, 1, 0)) 
```

```{r}
dat.philly <- dat.philly %>%
  mutate(geometry = st_geometry(PhillyFinal))
dat.philly <- st_as_sf(dat.philly, sf_column_name = "geometry")
```

Plot the Predictive Map
```{r}
ggplot() + 
  geom_sf(data=dat.philly, aes(fill = as.factor(gentrified))) + 
  geom_sf(data = PhillyTracts, fill = NA, color = "black") +
  theme_minimal() +
  labs(title="Predicted Gentrification Area in Philadelphia") +
  theme(plot.title = element_text(hjust = 0.5),  
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
    legend.position = "right")
```

## Calculating Gentrified Area in Philly 
```{r, eval=FALSE}

# Download data for 2009 and 2019
p.data_2009 <- get_acs(geography = "tract",
  variables = c(education2009, income_var, property_var),
  year = 2009,
  state = "PA",
  county = "Philadelphia",
  geometry = TRUE,
  output = "wide")


p.data_2019 <- get_acs(
  geography = "tract",
  variables = c(education_vars, income_var, property_var),
  year = 2019,
  state = "PA",
  county = "Philadelphia",
  geometry = TRUE,
  output = "wide")


```
```{r}
process_education_data2009 <- function(data) {
  data %>%
    mutate(college_degree_or_higher = rowSums(select(., B15002_014E, B15002_015E,
                                                      B15002_016E, B15002_017E, 
                                                      B15002_018E, B15002_031E,
                                                      B15002_032E, B15002_033E, 
                                                      B15002_034E, B15002_035E), na.rm = TRUE)) %>%
    select(GEOID, college_degree_or_higher, ends_with("B19013_001E"), ends_with("B25077_001E"))
}

# Processing function for 2019 data
process_education_data2019 <- function(data) {
  data %>%
    mutate(college_degree_or_higher = rowSums(select(., B15003_021E, B15003_022E, 
                                                     B15003_023E, 
                                                     B15003_024E, B15003_025E), 
                                              na.rm = TRUE)) %>%
    select(GEOID, college_degree_or_higher, ends_with("B19013_001E"), ends_with("B25077_001E"))
}

```

```{r, include=FALSE}

# Download data for 2009 and 2019
p.data_2009 <- get_acs(geography = "tract",
  variables = c(education2009, income_var, property_var),
  year = 2009,
  state = "PA",
  county = "Philadelphia",
  geometry = TRUE,
  output = "wide")


p.data_2019 <- get_acs(
  geography = "tract",
  variables = c(education_vars, income_var, property_var),
  year = 2019,
  state = "PA",
  county = "Philadelphia",
  geometry = TRUE,
  output = "wide")


```
```{r}
process_education_data2009 <- function(data) {
  data %>%
    mutate(college_degree_or_higher = rowSums(select(., B15002_014E, B15002_015E,
                                                      B15002_016E, B15002_017E, 
                                                      B15002_018E, B15002_031E,
                                                      B15002_032E, B15002_033E, 
                                                      B15002_034E, B15002_035E), na.rm = TRUE)) %>%
    select(GEOID, college_degree_or_higher, ends_with("B19013_001E"), ends_with("B25077_001E"))
}

# Processing function for 2019 data
process_education_data2019 <- function(data) {
  data %>%
    mutate(college_degree_or_higher = rowSums(select(., B15003_021E, B15003_022E, 
                                                     B15003_023E, 
                                                     B15003_024E, B15003_025E), 
                                              na.rm = TRUE)) %>%
    select(GEOID, college_degree_or_higher, ends_with("B19013_001E"), ends_with("B25077_001E"))
}

```

```{r}
p.data_2009 <- p.data_2009 %>%
  st_set_geometry(NULL)

p.data_2019 <- p.data_2019 %>%
  st_set_geometry(NULL)

p.data_2009_processed <- process_education_data2009(p.data_2009)

p.data_2009_processed <- p.data_2009_processed %>%
    mutate(income_group = cut(B19013_001E,
                            breaks = quantile(B19013_001E, probs = c(0, 1/3, 2/3, 1), 
                                              na.rm = TRUE),
                            labels = c("Low", "Medium", "High"),
                            include.lowest = TRUE)) %>%
  filter(income_group %in% c("Low", "Medium")) %>%
  na.omit()

p.data_2019_processed <- process_education_data2019(p.data_2019)


p.combined_data <- left_join(p.data_2009_processed, p.data_2019_processed, by = "GEOID")


p.combined_data <- p.combined_data %>%
  mutate(
    percent_change_education = (college_degree_or_higher.y - college_degree_or_higher.x) / college_degree_or_higher.x * 100,
    percent_change_income = (B19013_001E.y - B19013_001E.x) / B19013_001E.x * 100,
    percent_change_property = (B25077_001E.y - B25077_001E.x) / B25077_001E.x * 100
  ) %>%
  na.omit()

```

Identify Gentrified Areas

```{r}
mean(p.combined_data$percent_change_income) 
mean(p.combined_data$percent_change_property) 
mean(p.combined_data$percent_change_education) 

p.combined_data <- p.combined_data %>%
  mutate(actualgentrified = ifelse(percent_change_income > 33.42189 & 
                               percent_change_property > 43.26612 & 
                               percent_change_education > 73.71237, 1, 0)) %>%
  select(actualgentrified, GEOID)

dat.3 <- dat.philly %>%
  st_join(PhillyFinal)

dat.3 <- dat.3 %>%
  group_by(GEOID) %>%
  slice(1)

dat.3 <- dat.3 %>%
  left_join(p.combined_data, by = "GEOID")

dat.3 <- dat.3 %>%
  na.omit()

```

*Comparing Predicted Map with Observed Map*

We gathered the same variables from Philadelphia in 2019 and incorporated them into our prediction model, resulting in the predicted gentrification map displayed in below. Upon comparing this with the observed gentrification map in Philly, it becomes evident that the data from Chicago and our model could be effectively adapted to predict gentrification trends in Philadelphia. This provides valuable insights for planners, demonstrating the model’s applicability and effectiveness in different urban contexts. 


```{r}
ggplot() + 
  geom_sf(data=dat.3, aes(fill = as.factor(gentrified))) + 
  geom_sf(data = PhillyTracts, fill = NA, color = "black") +
  theme_minimal() +
  labs(title="Predicted Gentrification Area in Philadelphia") +
  theme(plot.title = element_text(hjust = 0.5),  
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
    legend.position = "right")
```


```{r}
ggplot() + 
  geom_sf(data=dat.3, aes(fill = as.factor(actualgentrified))) + 
  geom_sf(data = PhillyTracts, fill = NA, color = "black") +
  theme_minimal() +
  labs(title="Actual Gentrification Area in Philadelphia") +
  theme(plot.title = element_text(hjust = 0.5),  
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
    legend.position = "right")

```











